---
title: "JOUR328T_ggplot"
output: html_document
date: "2024-03-18"
---



#### Load packages
```{r}
library(tidyverse)
library(scales)
library(treemapify)
library(RColorBrewer)
```


#### Load Spotify data csv
```{r}
spotify_data <- read.csv("Most_Streamed_Spotify _Songs_2023.csv")
```

We'll start by getting our data ready.
```{r}
spotify_bar_chart_data <- spotify_data %>% top_n(10, streams)
```

Let's make the chart and order it by number of streams
```{r}
spotify_bar_chart_data %>% ggplot(aes(x = reorder(track_name,desc(streams)), y = streams)) +
  geom_col()
```

Next, fix the tick labels
- Set y axis to display in billions
- Set x on angle
(To do -- add theme, headline, chatter, other visual improvements)

```{r}
spotify_bar_chart_data %>% ggplot(aes(x = reorder(track_name,desc(streams)), y = streams)) +
  geom_col()+
  scale_y_continuous(labels = scales::label_number_si()) +
  scale_x_discrete(guide = guide_axis(angle = 45))
```
Add some design elements 
- Make the bars dark green
- Add text
- Add a theme (minimal)
- hide the axis labels

```{r}
spotify_bar_chart_data %>% ggplot(aes(x = reorder(track_name,desc(streams)), y = streams)) +
  geom_col(fill="darkgreen")+
  scale_y_continuous(labels = scales::label_number_si()) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(title = "Top Spotify Tracks of 2023", subtitle = "The Weekend's Blinding Lights takes the top spot with over 3 billion streams.", caption = "Source: Spotify via Kaggle \n by Adam Marton")+
  theme_minimal()+
  theme(axis.title = element_blank())

```



Next, let's make a scatterplot based on a few values. I was curious if energy and danceability are related. Are they? 
(To do: break this up in a few steps -- scatterplot, then add color and size -- add theme, headline and other visual elements)

```{r}
  ggplot(spotify_data, aes(danceability_., energy_.)) +
    geom_point()

```

Let's add in color based on the key column

```{r}
  ggplot(spotify_data, aes(danceability_., energy_., color = key)) +
    geom_point()
```


Let's encode another variable -- speechiness -- by size. And we will make the dots a little transparent and use RColorBrewer to choose our own colors
```{r}
  ggplot(spotify_data, aes(danceability_., energy_., color = key)) +
    geom_point(alpha = .5, aes(size = speechiness_.))+
    theme_minimal() +
    scale_color_brewer(palette="Paired")
```


Add text and clean up labels (axis and legend labels)
```{r}
  ggplot(spotify_data, aes(danceability_., energy_., color = key)) +
    geom_point(alpha = .5, aes(size = speechiness_.))+
    theme_minimal() +
    scale_color_brewer(palette="Paired") +
   labs(title = "Are energy and danceability related?", subtitle = "Examining the top Spotify songs of 2023", caption = "Source: Spotify via Kaggle \n by Adam Marton", x = "Danceability", y = "Energy", size = "Speechiness", color = "Key")
```



Use facet wrap to make it into small multiples
```{r}
  ggplot(spotify_data, aes(danceability_., energy_., color = key)) +
    geom_point(alpha = .5, aes(size = speechiness_.))+
    theme_minimal() +
    scale_color_brewer(palette="Paired") +
   labs(title = "Are energy and danceability related?", subtitle = "Examining the top Spotify songs of 2023", caption = "Source: Spotify via Kaggle \n by Adam Marton", x = "Danceability", y = "Energy", size = "Speechiness", color = "Key") +
  facet_wrap(~key)
```





Let's prep our data for making an area by creating a pivot table for year
```{r}
spotify_area_data <- spotify_data %>% 
  group_by(released_year) %>%
  summarize(count_by_year = n())
```


Let's make an area chart to see the spread of when songs were released
(To do: clean up, add theme, add hed and chatter)
```{r}
ggplot(spotify_area_data, aes(x=released_year, y=count_by_year)) +
  geom_area(alpha = 0.5, fill="darkgreen")
```
Add headline and chatter and theme

```{r}
ggplot(spotify_area_data, aes(x=released_year, y=count_by_year)) +
  geom_area(alpha = 0.5, fill="darkgreen") +
  labs(title = "Most of Spotify's most popular songs of 2023 were recently released", subtitle = "Though songs go back as far as 1930", caption = "Source: Spotify via Kaggle \n by Adam Marton", x = "Year", y = "Count") +
    theme_minimal() 
```




Let's prep our data for making a tree map by creating a pivot table for KEY
```{r}
spotify_tree_data <- spotify_data %>% 
  group_by(key) %>%
  summarize(count_by_key = n())
```



Now lets make a tree map showing songs by key
Here is a guide to treemapify: https://r-charts.com/part-whole/treemapify/
(to do -- add labels, theme, headline, what else from link above?)
```{r}
ggplot(spotify_tree_data, aes(area = count_by_key, fill = key)) +
  geom_treemap() +
      theme_minimal() +
   labs(title = "Top Spotify songs by key", subtitle = "Examining the top Spotify songs of 2023", caption = "Source: Spotify via Kaggle \n by Adam Marton", color = "Key")
```

Add labels, theme, headline, custom colors
```{r}
ggplot(spotify_tree_data, aes(area = count_by_key, fill = key, label = count_by_key)) +
  geom_treemap() +
      theme_minimal() +
   labs(title = "Top Spotify songs by key", subtitle = "Examining the top Spotify songs of 2023", caption = "Source: Spotify via Kaggle \n by Adam Marton", color = "Key")+
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 15)+
    scale_fill_brewer(palette = "Paired")

```

